---
title: "BIOL432 Assignment 3"
output: html_document
---

# Project Info
Created by: Caroline Tang (20115082)  
Github Username: [carolinetang77](https://github.com/carolinetang77/)  
Date: Jan 25, 2022  
[Project link](https://github.com/carolinetang77/biol432-assignment3)  

# Data Wrangling
```{r setup}
library(dplyr)
```

Reading in the data
```{r fallopia wrangling}
dat <- read.csv("./InputData/FallopiaData.csv")
```

Filtering, Selecting, Adding a new column
```{r}
wrangled <- dat %>% 
  #remove rows with Total < 60
  filter(Total >= 60) %>% 
  #select Total, Taxon, Scenario, Nutrients in that order and remove all other columns
  select(Total, Taxon, Scenario, Nutrients) %>% 
  #replace total (currently in mg) with total in grams
  mutate(Total = Total/1000) %>% 
  #name this new column TotalG
  rename(TotalG = Total)
print(wrangled)
```

# Custom Function
```{r}
calculate <- function(x, calculation) {
  if (calculation == "Average") { #if string is "Average" calculate the mean
    cat("Calculating mean...\n")
    result <- mean(x, na.rm = T)
    
  } else if (calculation == "Sum") { #if string is "Sum" calculate the sum
    cat("Calculating sum...\n")
    result <- sum(x, na.rm = T)
    
  } else if (calculation == "Observations") { #if string is "Observations" count number of observations
    cat("Counting up observations...\n")
    result <- length(x)
  } else {
    stop(message("Calculation not recognized"))
  }
  return(result)
}
```
### Testing the function

Counting observations in Taxon column
```{r}
calculate(wrangled$Taxon, "Observations")
```
Calculating average TotalG for the high nutrient group
```{r}
calculate(wrangled$TotalG[wrangled$Nutrients == "high"], "Average")
```
Calculating average TotalG for the low nutrient group
```{r}
calculate(wrangled$TotalG[wrangled$Nutrients == "low"], "Average")
```

# Saving the new data
```{r}
write.csv(wrangled, file = "./Output/WrangledData.csv", row.names = F)
```

